---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
```{r}
#wd
setwd("C:/Users/JARVIS/Desktop/Uni/Semester 4/Computational modelling/Portfolio-2-Cog-sci-knowledge")

#libraries
pacman::p_load(readr,groupdata2,ggplot2,tidyverse,data.table, rethinking)

#define grid
dens=1000
p_grid<- seq(from=0, to=1, length.out=dens)

#define prior, flat prior in this case
prior<- rep(1,dens)
#prior<- ifelse(p_grid<0.5, 0, 1) informed 
#prior<- exp(-5*abs(p_grid-0.5)) conservative
```

Riccardo:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(3, size = 6, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dR <- data.frame(
  Name=rep("Riccardo",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

## GGPLOT, TO MAKE IT PRETTY
ggplot(dR, aes(Grid,Posterior))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Riccardo's knowledge estimated with a flat prior")

#plot the prior too
ggplot(dR, aes(Grid,Prior))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Prior probability") +
  labs(title="The flat prior")

#QUADRATIC APPROXIMATION
#riccardo.qa<- map( #provide a formula, a list of data, a list of start values for the parameters
  #alist(w~dbinom(q,p), #binomial likelihood
        #p~dunif(0,1) #uniform prior, uniform distribution
        #),
  #data=list(w=3, q=6))

riccardo.qa<- map( #provide a formula, a list of data, a list of start values for the parameters
  alist(w~dbinom(6,p), #binomial likelihood
        p~dunif(0,1) #uniform prior
        ),
  data=list(w=3))

#display summary of qudratic approximation
precis(riccardo.qa) 

#mean - peak at 0.5
#stddev - 0.2

#quadratic aaproximation
curve(dnorm(x,0.5,0.2)/2, lty=2, xlab="probability of Riccardo giving a correct answer", ylab="posterior probability")
mtext("Posterior: quadratic approximation")


#What's his most likely parameter for knowledge?
max(posterior)
p_grid[posterior==max(posterior)] #0.4994995 - 0.5005005 

#What's his estimated knowledge? - THE ENTIRE POSTERIOR!
 
#What's the probability of him knowing better than chance?
#add up posterior probability where p>0.5
sum(posterior[p_grid>0.5])
```



2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

Kristian:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(2, size = 2, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dK <- data.frame(
  Name=rep("Kristian",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

## GGPLOT, TO MAKE IT PRETTY
ggplot(dK, aes(Grid,Posterior))+
  geom_line()+
  theme_classic()+
  
  xlab("knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Kristian's knowledge estimated with a flat prior")


#add up posterior probability where p>0.5
sum(posterior[p_grid>0.5])
```

Josh:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(160, size = 198, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dJ <- data.frame(
  Name=rep("Josh",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

## GGPLOT, TO MAKE IT PRETTY
ggplot(dJ, aes(Grid,Posterior))+
  geom_line()+
  theme_classic()+
  xlab("knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Josh's knowledge estimated with a flat prior")


#add up posterior probability where p>0.5
sum(posterior[p_grid>0.5])
```

Mikkel:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(66, size = 132, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dM <- data.frame(
  Name=rep("Mikkel",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

## GGPLOT, TO MAKE IT PRETTY
ggplot(dM, aes(Grid,Posterior))+
  geom_line()+
  theme_classic()+
  xlab("knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Mikkel's knowledge estimated with a flat prior")


#add up posterior probability where p>0.5
sum(posterior[p_grid>0.5])
```

Together, to actually compare:
```{r}
#add up dataframes
total <- rbind(dR, dK, dJ, dM)

#plot posteriors together
ggplot(total, aes(x=Grid, y=Posterior, color=Name))+
  geom_line()+
  theme_classic()+
  xlab("knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Teachers' knowledge estimated with a flat prior")

```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.
The prior:
```{r}
#define grid
dens=1000
p_grid<- seq(from=0, to=1, length.out=dens)

#define prior, prior is a normal distribution 
prior<- dnorm(p_grid, mean=0.8, sd=0.2)
```

Riccardo:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(3, size = 6, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dRn <- data.frame(
  Name=rep("Riccardo",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


## GGPLOT, TO MAKE IT PRETTY
ggplot(dRn, aes(Grid,Posterior))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Riccardo's knowledge estimated with a normal prior")

##plot old posterior too
#add up dataframes
totalR <- rbind(dR,dRn)
ggplot(totalR, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Riccardo's knowledge estimated with a uniform vs normal prior")

#plot the prior too
ggplot(dRn, aes(Grid,Prior))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Prior probability") +
  labs(title="The normal prior")

```

Repeat for Kristian:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(2, size = 2, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dKn <- data.frame(
  Name=rep("Kristian",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


##plot old posterior too
#add up dataframes
totalK <- rbind(dK,dKn)

#plot uniform vs normal
ggplot(totalK, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Kristian's knowledge estimated with a uniform vs normal prior")


```

Now Josh:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(160, size = 198, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dJn <- data.frame(
  Name=rep("Josh",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


##plot old posterior too
#add up dataframes
totalJ <- rbind(dJ,dJn)

#plot uniform vs normal
ggplot(totalJ, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Josh's knowledge estimated with a uniform vs normal prior")
```
Now Mikkel:
```{r}
#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(66, size = 132, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dMn <- data.frame(
  Name=rep("Mikkel",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


##plot old posterior too
#add up dataframes
totalM <- rbind(dM,dMn)

#plot uniform vs normal
ggplot(totalM, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Mikkel's knowledge estimated with a uniform vs normal prior")
```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
dR<- NULL
dRn <- NULL
totalR<- NULL
dK <- NULL
dKn <- NULL
totalK<- NULL
dJ<- NULL
dJn<- NULL
totalJ <- NULL
dM <- NULL
dMn <- NULL
totalM <- NULL

```


Riccardo:
```{r}
#define prior, flat prior in this case
prior<- rep(1,dens)

#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(300, size = 600, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dR <- data.frame(
  Name=rep("Riccardo",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

######## NORMAL PRIOR
#define prior, prior is a normal distribution 
prior<- dnorm(p_grid, mean=0.8, sd=0.2)

#likelihood is the same

#compute product of likelihood and normal prior
unstd.posterior<- likelihood*prior

#standardize the posterior for flat prior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dRn <- data.frame(
  Name=rep("Riccardo",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))

##plot old posterior too
#add up dataframes
totalR <- rbind(dR,dRn)
ggplot(totalR, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Riccardo's knowledge estimated with a uniform vs normal prior")
```



Kristian:
```{r}
#define grid for Kristian
dens=1000
p_grid<- seq(from=0, to=1, length.out=dens)

#define prior, flat prior in this case
prior<- rep(1,dens)

#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(200, size = 200, prob = p_grid)

#compute product of likelihood and flat prior
unstd.posterior<- likelihood*prior

#standardize the posterior for flat prior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dK <- data.frame(
  Name=rep("Kristian",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

######## NORMAL PRIOR

#define normal distribution prior
prior<- dnorm(p_grid, mean=0.8, sd=0.2)

#likelihood is the same

#compute product of likelihood and normal prior
unstd.posterior<- likelihood*prior

#standardize the posterior for flat prior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dKn <- data.frame(
  Name=rep("Kristian",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


##plot old posterior too
#add up dataframes
totalK <- rbind(dK,dKn)

#plot uniform vs normal
ggplot(totalK, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Kristian's knowledge estimated with a uniform vs normal prior")
```


Josh:
```{r}
#define grid
dens=1000
p_grid<- seq(from=0, to=1, length.out=dens)

#define prior, flat prior in this case
prior<- rep(1,dens)

#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(16000, size = 19800, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dJ <- data.frame(
  Name=rep("Josh",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))

## GGPLOT, TO MAKE IT PRETTY
ggplot(dJ, aes(Grid,Posterior))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Josh's knowledge estimated with a flat prior")

######## NORMAL PRIOR
#define prior, prior is a normal distribution 
prior<- dnorm(p_grid, mean=0.8, sd=0.2)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dJn <- data.frame(
  Name=rep("Josh",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


##plot old posterior too
#add up dataframes
totalJ <- rbind(dJ,dJn)

#plot uniform vs normal
ggplot(totalJ, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Josh's knowledge estimated with a uniform vs normal prior")
```

Mikkel:
```{r}
#define grid for Mikkel
dens=1000
p_grid<- seq(from=0, to=1, length.out=dens)

#define prior, flat prior in this case
prior<- rep(1,dens)


#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(6600, size = 13200, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dM <- data.frame(
  Name=rep("Mikkel",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Uniform",dens))


######### Normal prior 
#define prior, prior is a normal distribution 
prior<- dnorm(p_grid, mean=0.8, sd=0.2)

#likelihood is the same

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dMn <- data.frame(
  Name=rep("Mikkel",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))


##plot old posterior too
#add up dataframes
totalM <- rbind(dM,dMn)

#plot uniform vs normal
ggplot(totalM, aes(x=Grid, y=Posterior, color=Priortype))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Posterior probability") +
  labs(title="Mikkel's knowledge estimated with a uniform vs normal prior")
```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
```{r}
#normal distribution with the mean 0.5 and std dev of 0.1
prior<- dnorm(p_grid, mean=0.5, sd=0.1)

#compute likelihood at each value in grid// binomial distribution function (how many correct, size = how many trials, what would be the probability of this data) <- gonna get probability for each possible rate of the correct answer
likelihood<- dbinom(160, size = 198, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior<- likelihood*prior

#standardize the posterior, so it sums to 1
posterior<- unstd.posterior/sum(unstd.posterior)

dRn <- data.frame(
  Name=rep("Riccardo",dens),
  Grid = p_grid,
  Prior= prior,
  Posterior = posterior,
  Priortype=rep("Normal",dens))

#plot the prior too
ggplot(dRn, aes(Grid,Prior))+
  geom_line()+
  theme_classic()+
  xlab("Knowledge parameter")+ ylab("Prior probability") +
  labs(title="The skeptic prior")

```


6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



